Responses to Reviews

### 2016-07-05 RMS/responses.tex
updated response.tex, ready for 2nd rev

\begin{center}
  \begin{tabular}{r l}
    \hline\hline
    & \\
    Manuscript Number: & 15-TIE-3480\\
    Manuscript Title: & Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation\\
    & with Real-Time Maude \\
    Submitted to: & Transactions on Industrial Electronics \\
    Manuscript Type: & Regular paper \\
    & \\
    \hline\hline
  \end{tabular}
  \bigskip
\end{center}

@@

\section{Responses to Reviewer 1}
\subsection{General Comments}

\hide{ The paper shows a realistic implementation of a Rate-Monotonic
  Scheduling algorithm using Real-Time Maude (a modeling language
  based on rewriting logic) by taking into account the overhead of
  scheduling and some other details of the hardware platform. The
  correctness of the implementation with respect to the algorithm and
  the completeness of the model are verified by model checking, and
  validated within different realistic scenarios.}

@@

On the positive side, this paper presents a solid piece of
work. Periodic task scheduling is an important (and difficult)
problem, and the topic of the paper appears to be well suited within
the field of the industrial real-time systems. Overall the paper is
well structured, technically accurate, presents the work in a
comprehensive and reasonable way, and tackles theoretical as well as
practical aspects. There are two principal contributions of the paper:

1) A novel implementation of the Rate-Monotonic Scheduling algorithm,
which contains more complex and realistic details instead of the ideal
setting.

2) Rate-Monotonic Scheduling is investigated using Real-Time Maude for
the first time, to apply formal methods such as model checking and
theorem proving to analyze theoretical results, verify desired
properties, and evaluate results.

@@

Thanks for the pros. We are happy that the work in the paper interests
the readers. We hope that it provides a new and formal way to verify a
real-time system, which is able to investigate important problems
(such as periodic task scheduling) in a more realistic level.

@@

The weak point is the lack of evidence of practicality, efficiency and
scalability. Since there are only small and simplistic scenarios to
analyze this model, it is difficult to judge the effectiveness of the
proposed implementation. The key scenarios appear conveniently small
to allow the implementation fully supports all these examples,
returning at the same time appropriate counterexamples to guide and
adjust the design. Is this true for larger examples of "real world"
applications within the field of industrial real-time systems? How do
the authors propose to handle very large scenarios?  How does it
behave on typical real-time scheduling problems used in industry? I
would like to see more complex examples to convince the reader of the
practicability of the approach on the assumptions given for the model,
and would probably increase the mentioned benefits. The paper does not
give any directions to the above problems, and some sentences pointing
out the difficulties in the implementation of more generic and
realistic scenarios would be most welcome to understand the real
contribution of the paper.

@@

Thanks for the suggestions. As mentioned in the paper, the system
under verification is a real-world RMS implementation, which is used
in an avionic control system. The scenarios presented in Section V.B
are also real scenarios from the online system. Our industrial partner
has very strict timing requirements. They use at most 5 tasks for
scheduling to avoid high overhead. They agree that the verification
results presented in the paper satisfy their practical requirements.

On the other hand, as presented in the revised Section V.B, we have
examined our approach by verifying randomly generated scenarios where
the number $n$ of tasks are over $10$. In that case, the least common
multiple of the periods of all tasks can be as large as $10^5$ times
the interrupt cycle $T$. All feasible combinations of execution traces
are checked in our approach. Furthermore, the number $n$ of tasks can
be $20$ and even more if we allow different tasks possess a same
period.

\hide{
the modeled RMS implementation serves in an avionic control systems
from our industrial partner. It is a real-world application within the
field of industrial real-time systems. And the scenarios presented in
Section V.B are also real scenarios that our partner is using in the
system. Our partner indicated that there would be at most $5$ tasks in
the task set for the implementation. On the other hand, in the current
version, we have randomly generated large numbers of test cases to
examine the efficiency of our schedulability test that is based on
model checking, as discussed in Section V.B. The capability of model
checking is restricted by the scale of the state space of the
model. In our model, the model checking can handle scenarios where
$mn$ is up to $10^6$ in an acceptable period of time. With $mn=10^6$
and reasonable task periods, the number $n$ of tasks can be at least
$10$, which has already fulfilled the requirements of the target
implementation. Further with particular task periods, the number $n$
of tasks can be $20$ or more.
}

One of my biggest problems with the theoretical part of this work is
that the paper is technically solid as far as I can tell, but there's
not much yet in the way of theorems and results about formal
properties in Maude. The complete and detailed proof of the Theorem 2
should be provided somewhere. If the proof of the main results of
schedulability and correctness of the model is straightforward (as is
suggested) it seems not necessary to present a complex formalism on
Real-Time Maude and all the theorems provided are natural consequences
of Theorem 1. I agree that even non-surprising results need to be
investigated in the proof of Theorem 2, some sentences pointing out
the difficulties in the theoretical formalization in Real-Time Maude
would be most welcome. Please, include (e.g., in an appendix or even
in a technical report) a detailed proof.

We are happy that the readers are interested in the detailed technical
proof, and we have now included it in the Appendix. It is true that
readers may be interested in the detailed proof of Theorem~2 to feel
convinced of the completeness of our approach. However, the proof
needs a bit more theoretical background about rewriting logic and more
technical details of our model, which we have simplified in the paper
in order to ease the understanding of our approach. We would prefer to
refer the detailed proof to a non-anonymous technical report, if
possible, in the final version.

Another important point which needs improvement is the description of
the implementation. I guess other readers might have less difficulties
with the motivations if they know Maude and the Real-Time Maude
extension, but if you aim for a more general audience, it might not
work. Many pages are spent for the description of the formal modeling
of the implementation and the section is hard to read. I highly
recommend the authors to thoroughly revise this part of the paper to
make it easy-to-read. It would be better preferable to have a much
shorter description of the principal insights for the implementation
followed by more convincing examples and benchmarks with respect to
related work to convince the reader of the practicability of the
approach. You should try to summarize this section and to extend the
original contribution. From my point of view, there is a trade off in
this paper between the contributions and the presentation of the
implementation.

Thank you for the invaluable suggestions. Section IV has been
revised. Some unnecessary detailed definitions presented as code are
removed, and replaced by more literal explanations. We hope that the
current presentation would make Section IV more easy-to-read, even for
an audience who did not know Real-Time Maude before. On the other
hand, Section VI (Related Work) is also improved. We compare our work
with the existing theoretical work and the existing verification work
in a more complete way.

\subsection{Detailed Comments}

Page 7, Related Work. More comparisons with other approaches based on
model checking would help to evaluate pros and cons of an
implementation with Real-Time Maude with respect to different
languages and tools. The authors should review the comparison with
[24] and [25] more thoroughly. You should benchmark your approach
against others.

Thanks for the pointing this out. The related work section is now
improved. More comparisons with the theoretical approaches and with
[24,25] are added. The differences with [24,25] in the models and in
the way of modeling are mainly discussed. On the other hand, it is a
bit difficult to compare the efficiency and scalability between
[24,25] and our approach. One reason is that we have a different
setting with [24] and [25], since the objectives are different: we aim
at verifying an RMS implementation from a real-world real-time system,
while [24] and [25] targeted the RMS algorithm itself. Another reason
is that [24] and [25] presented no discussion on the efficiency and
scalability. Furthermore, they used \verb|TMSVL| and an extension of
SPIN, respectively, as verification tools, which are not open-source,
making us unable to re-implement their work.

Page 8, Conclusions. Where in the paper do you show that the details
of your "realistic" implementation are "sufficient" for the behaviors
of all real systems used in the current industry? This should be
briefly discussed.

Sorry for the confusion. The statement has been corrected. The work
presented in the paper aims at modeling and verifying a realistic RMS
implementation in an industrial avionic system. It is a piece of real
work in the industry. In this sense, ``sufficient'' means that the
details described in our model meet the requirements and expectations
from the users and the engineers. On the other hand, we believe that
our approach could be applied to other similar systems as well.

\section{Responses to Reviewer 2}
\subsection{General Comments}

One of the problems with this paper is that it is not completely
self-contained.  Some knowledge about Real-Time Maude is necessary in
order to fully understand it.  I found the short introduction about
Real-Time Maude (section II.B) not enough for understanding the
formalism used in the rest of the paper (see details below).

Thank you very much for pointing this out. The Sections II.B and IV
are improved to be more self-contained now. The idea is that, Section
II.B presents mainly an overview of Real-Time Maude, and then Section
IV describes the model with introducing necessary syntax on-the-fly.

Another concern is that the significance of the paper could be much
better if the paper addressed the problem of modelling RMS in a more
general way. Instead, the paper takes the form of a case study, where
the formal modelling of one specific RMS implementation is presented,
rather than a general method for modelling and analyzing RMS
implementations.

Thank you for the suggestion. Yes, we agree that we have two choices
to do the work presented in the paper: one is to develop a general way
to model RMS algorithms or implementations, and then instantiate the
general model to get a concrete one for our target system; the other
is to shape an accurate model of the target system directly.

We chose the latter according to the following considerations. By
experience, \emph{the development engineers (who write the code) and
  verification engineers (who verify the code or the system) are
  usually not the same persons.} When a verification engineer models a
system, he is actually abstracting the system or the code. It is
difficult to ensure that the model behaves the same as the system or
the code.  And it is more difficult to make other engineers and users
believe that the model behaves the same as the system or the
code. Furthermore, the developers and users hope that the model
behaves not only the same as what they had in mind, but also the same
as what is written in the code. A way out is to make the model not
only \emph{behave like} the code, but also \emph{``look'' like} the
code. This requires the model to be as accurate as possible, based on
a fact that the expressiveness of the modeling language is powerful
enough. It makes the model specific, but improves the engineers'
confidence in the model and the verification results. That is also why
we emphasize the corresponding relationship between the functions in
the model and the lines in the pseudocode, when we introduce our model
in Section~IV.

On the other hand, we are also interested in the first direction. This
will be the future work and we would try to get a better balance
between them.

Also, the discussion about soundness and completeness of the analysis
(section V.C) is not fully developed and, once again, it refers to the
specific model rather than being a general result.

Thanks for this comment. The detailed proof of Theorem~2 is now
included in the Appendix. On the other hand, since we chose the
accurate way to achieve the model as discussed above, yes, the proof
is a specific one.

\subsection{Detailed Comments}

[Section II.B] In the definition of $IR$, it is not clear what $s$ and
$s'$ are. I guess they are states, but the concept of state was not
introduced.

Yes, they can be states. But more generally, they are terms. This part
has been modified to specify $s$ and $s'$ (in fact, $t$ and $t'$ now)
clearly.

[Section II.B] In the definition of a class, the authors should
specify that a class includes a collection of rules, otherwise this
fact may be missed.

Thanks very much for reminding. This is added now.

[Section III] Page 2 right column, line -24: The implementation is
shown as schedule() in Figure 1 $\rightarrow$ The pseudocode of
schedule() is shown in Figure 1

Thanks for the careful reading. It is done now.

[Section IV.A] I don't understand the meaning of \verb|[ctor]| in the
definition of some operations (maybe because I have no specific
background on Maude). I could not find the explanation of this writing
in the introduction to Maude.

Sorry for missing the explanation of \verb|[ctor]|. It is now added
when \verb|[ctor]| is used for the first time.

[Section IV.A] For the stack sort, the authors mention operations
push, pop and peek, but then you don't write the definitions of these
operations. Instead, they write the definitions of bottom and \#,
which are not explained in the text. This part should be made clearer.

Sorry for the misleading presentation in the text. In fact,
\verb|bottom| and \verb|#| are constructors of sort \verb|Stack|. This
part is modified now. Some detailed definitions in code are removed,
and are replaced by more explanations. We hope that the current
presentation would make Section IV more easy-to-read.

[Section IV.A] Also, the meaning of NzNat is not explained.

Sorry for this incompleteness. \verb|NzNat| meant non-zero natural
numbers. \verb|NzNat| is now replaced by \verb|Nat| to reduce
unnecessary complexity for the readers.

[Section V.C] In the statement of Theorem~1, the definition of
time-robust real-time rewrite theory is missing. Also, the meaning of
tick-stabilizing atomic propositions is not defined. This makes the
section not self-contained.

Sorry for the inconvenience for reading. In fact, the definitions of
time-robustness and tick-stabilization (which is now replaced with
tick-invariance) require more knowledge about rewriting logic. We
avoid introducing the accurate definitions of them. Instead, now we
give short descriptions of them before Theorem~1. More detailed
introduction can be found in the Appendix.

[Section V.C] The proof of theorem 2 is missing. The authors just
indicate how the proof could be developed, but they do not develop
it. A possibility would be to point to a document where this proof has
been developed.

We are very happy that the readers are interested in the detailed and
technical proof, which is now included in the Appendix.  The detailed
proof requires a bit more theoretical background about rewriting logic
and more details of our model, which we tried to simplify in the paper
in order to ease the understanding of our approach. We would prefer to
refer the detailed proof to a non-anonymous technical report, if
possible, in the final version.

@@

RMS has many different realization methods, no comparison is given to
justify the strong point of real-time maude when applied to RMS.

@@

The comparison could be found in Section VI (and we have expanded
according to the reviewers' comments :-)). We would like to clarify
that we are \emph{not implementing} RMS, we are actually
\emph{verifying} it. We choose Real-Time Maude since it verifies the
model using model checking technique. If verification passes, no
deadline will be missed in the implementation as long as the model
assumptions are met.

\hide{
The comparison is given in the related work section (Section
VI). We would say that we are not implementing RMS algorithm in
Real-Time Maude. In fact, we are using Real-Time Maude to model an
realistic implementation of RMS from an industrial avionic
system. There is no theoretical reason telling us that we have to use
Real-Time Maude to model the RMS implementation. However, the existing
verification work as far as we know cannot fit our needs, as discussed
in Section VI. Therefore, we developed the work in the paper, as a
piece of real work in the industry.
}

Rate-monotonic scheduling is a very simple scheduling and has been
sufficiently investigated in the field of embedded system, the authors
claim the innovation of this paper as the implementation using
real-time maude, however, I cannot see any special point (or
advantage) when using this real-time maude in modelling RMS.

Thank you for the insightful comment. We agree that RMS has been
intensively studied because of its importance. However, most of the
results are based on the assumption that \emph{task switching does not
  take time}. That assumption is not realistic. Models with and
without the assumption have totally different behaviors. For instance,
as shown in Figure~2(b), in the RMS implementation we verified, the
switching from time $9$ to $11$ blocks the interrupt handling, hence
delaying the initiation time of $\tau_1$. This kind of phenomenon, to
our knowledge, would not happen in the theoretical analysis
models. Although there is very few results that take into account the
switching overhead (as discussed in Section VI), none of them is
generic and our system under verification is not in their scope.

Despite the above reason, the most important motivation of this paper
is to verify the system not only for the schedulability, but also for
the correctness. That is, we want to verify that the target system
does schedule the tasks under the RMS algorithm. The work presented
should be seen as a verification problem, instead of a schedulability
test. From verification point of view, this paper models a real-world
RMS implementation with sufficient technical details. Two important
properties--schedulability and correctness--are verified by model
checking technique, and the soundness and completeness of the results
are demonstrated.

\hide{
is a very simple scheduling and is the optimum at the same time among
the fixed-priority ones, making it widely applied and sufficiently
investigated. Besides the ideal setting, the schedulability of RMS was
also investigated under different kinds of implementations with
consideration about overhead, as we discuss in the related work
section (Section VI). Despite the reason that the existing results do
not fit our target system using RMS, the most important motivation to
develop the work presented is that, we aim at verifying the system not
only for the schedulability, but also for the correctness. That is, we
want to verify that the target system does schedule the tasks under
the RMS algorithm. The work presented should be seen as a verification
problem, instead of a schedulability test.
}

The assumptions given in IV is too rigid, which makes the scheduling
problem to be investigated by very simple, then what is the
difficulty?

Thank you for the question. When model checking technique is used to
verify a given system, the tool (model checker) explores the whole
system state space to check whether the given property holds. All
feasible combinations of execution traces are checked. As a result,
the state space explosion is the most common problem for a model
checker. That means the state space of the system/model is too huge
for the tool to explore with acceptable temporal and spatial cost. The
scale of the state space depends on the complexity of the model of the
system.

As an industrial verification application, we are not trying to make
the problem as difficult as possible. Instead, we are trying to
simplify the model so that the existing tools are able to solve, while
the target system satisfies the assumptions on the model so that the
verification results are trustworthy.

On the other hand, the assumptions are actually not that rigid. For
example, the mask mechanism considered makes the job initiation times
of task $\tau_i$ possibly not equal $kT_i$, implying that the
deadlines of the jobs may be not equal to $(k+1)T_i$ (see assumption
A1'). This is also discussed in the beginning of Section IV, pointing
out the blocking at time $10$ in Figure~2(b). Another example is that
non-determinism exists in our model, as discussed in Section~VI.

The description in section IV (A-E) is trivial, it can only be viewed
as a simple application of the rewriting logic.

Thanks for the careful reading. We agree that this is an application
of rewriting logic. In this paper, we do apply the existing theories
and techniques of rewriting logic to solve a real industrial problem,
which has not yet been investigated by this sort of techniques,
achieving an industrial contribution.

The model checking part is not sufficiently discussed, the authors
only list the commands but not illustrate the mechanism behind these
commands.

Thank you for the suggestion. Our work does focus on how to model a
realistic industrial system and how to apply formal techniques (such
as model checking) to verify important properties of the system. The
mechanism behind the model checking commands (i.e. of a model checker)
is another issue.

The field of model checking techniques and the implementation of model
checkers have been intensively studied for decades. If the reviewer is
interested in the mechanism of the model checker provided by Real-Time
Maude, we would recommend the reference [A] to the reviewer:

\noindent
[A] S. Eker, J. Meseguer, and A. Sridharanarayanan, ``The Maude LTL
Model Checker'', {\it Electr. Notes Theor. Comput. Sci.}, vol. 71,
pp. 162-187, 2002.

\hide{
However, since the model checker is not part of our contributions in
the paper, we do not think it is appropriate to illustrate the
mechanism behind the model checking commands, which may confuse the
readers on the other hand.}

>>>>

\begin{center}
  \begin{tabular}{r l}
    \hline\hline
    & \\
    Manuscript Number: & 15-TIE-3480.R1\\
    Manuscript Title: & Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation\\
    & with Real-Time Maude \\
    Submitted to: & Transactions on Industrial Electronics \\
    Manuscript Type: & Regular paper \\
    & \\
    \hline\hline
  \end{tabular}
  \bigskip
\end{center}

@@

\section{Responses to Reviewer 1}

@@

I have read the revised version of the article. Overall, the authors
did a very good job at addressing the reviewers' comments and I now
consider the paper to be acceptable for publication.

The revised version of this paper has been improved following the
suggestions given by the reviewers of the original submission:

(1) Sections II.B and IV are improved to be more self-contained now.

(2) A more detailed but simplified version of a proof of Theorem 2 is
now included in an Appendix, and the authors refer the detailed proof
to a non-anonymous technical report.

However, due to the importance of this theorem, I suggest to put the
complete detailed proof in the printed version.

In summary, I am satisfied with the revisions made by the authors and
consider that the paper can now be accepted for publication.

@@

\section{Responses to Reviewer 2}

@@

My most important concerns have been addressed, so I think the paper
is now in a shape acceptable for publication. I couldn't check the
proof completely, but it appears sound.

The significance of the paper remains not very high, because the paper
shows the proposed technique can be used in one particular case while
it does not provide results of general validity.

Nevertheless, as I said in my previous review, I believe these initial
results may be useful for readers, since there seem to be no other
paper providing a similar detail level in the models.

Detailed comments:

Section II.B, page 2

left column line 37-38

What are uj and vj? Are they terms too? This should be specified in
the text.

left column lines 42-46

I guess s' is the destination state. This should be specifioed in the
text.  The sentence "IR and TR together..." should not be part of the
item. It should go into a new paragraph after the list.

right column lines 22-24

As it seems the timed operators are not used, I suggest to avoid
mentioning them. By the way, this mention without any other detail is
useless.

Page 4, line 45

Why is a period of sort Nat? Shouldn't it be Counter?

@@

@@

This paper has been revised according to reviewers comments and
suggestions, I recommend to accept it for publication after minor
revisions on possible typos, such as the word "monus" on page 9 and
10.

@@
